### 2014-04-30 paper.tex
Some updates for wording and clarity for the gravitational boundary conditions.

\documentclass[12pt,preprint]{aastex}
% for \sout
\usepackage{ulem}
% makes sure \em{} is italic rather than underlined (corrects ulem from line above)
\normalem

% for the red MarginPars
\usepackage{color}

% some extra math symbols
\usepackage{mathtools}

\newcommand{\rhocutoff}{\rho_\mathrm{cutoff}}
\newcommand{\rhoanelastic}{\rho_\mathrm{anelastic}}

\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
\newcommand{\Tcutoff}{T_\mathrm{cutoff}}

% MarginPars
\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\newcommand{\evm}{{(-)}}
\newcommand{\evz}{{(\circ)}}
\newcommand{\evp}{{(+)}}
\newcommand{\enu}{{(\nu)}}

\newcommand{\msolar}{\mathrm{M}_\odot}

\begin{document}

%==========================================================================
% Title
%==========================================================================
\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code 
       Verification}

\shorttitle{DWD Mergers. I. Methodology}
\shortauthors{Max}

\author{TBD}
%==========================================================================
% Abstract
%==========================================================================
\begin{abstract}
We describe our numerical methodology for modeling double white dwarf
systems with the AMR hydrodynamics code Castro.

\end{abstract}
\keywords{hydrodynamics - methods: numerical - supernovae: general - white dwarfs}

%==========================================================================
% Introduction
%==========================================================================
\section{Introduction}

%==========================================================================
% Numerical Methodology
%==========================================================================
\section{Numerical Methodology}\label{sec:Numerical Methodology}

\subsection{Hydrodynamics}

We use the unsplit PPM solver in Castro to advance the hydrodynamics
system in time.  A number of changes were made to the solver, which are 
detailed in the Appendix.  These changes bring the algorithm more in 
line with that of \cite{ppm}.

The boundary conditions on the hyperbolic system are simply
zero-gradient.  Using AMR, we make the coarse grid very large, so as
to place the boundaries far from the region of interest.  We further
make the restriction that only level-0 (coarse) grids can touch the
domain boundary.

AMR

rotation

\subsection{Microphysics}

EOS (no Coulomb corrections)

\subsection{Initial Models}

We generate initial models by integrating the equation of hydrostatic
equilibrium, taking the temperature and composition to be constant,
and using the general stellar equation of state.  This results in
a single nonlinear equation to find the density in a zone given the
conditions in the zone beneath it:
\begin{equation}
\frac{p(\rho_{i+1}) - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}
\end{equation}
This is solved via Newton iteration to give $\rho_{i+1}$.  Here, $g_{i+1/2}$
is the gravitational acceleration at the interface between zones $i$ and $i+1$,
found by simply adding up all the mass from zones $1, \ldots, i$ to get the
enclosed mass, $M_{i+1/2}$, and then $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.

To start the integration off, we need a central density.  We guess at
this, and then iterate over the entire integration procedure to find
the central density needed to yield the desired total mass.  Finally,
we generate the initial model at the same resolution as the finest
grid in our simulation.

We map the 1-d model onto the 3-d Cartesian grid by taking density,
temperature, and composition as the independent variables,
interpolating these to the cell-centers, and then calling the equation
of state to initialize the remaining terms.  The velocity is taken to
be zero in the rotating frame initially.

\subsection{Gravity}

We solve the Poisson equation for self-gravity for our problem,
\begin{equation}
  \nabla^2 \Phi(\mathbf{x}) = -4\pi G\, \rho(\mathbf{x}),
\end{equation}
where $\Phi$ is the gravitational potential, $G$ is the gravitational constant, and $\rho$ is the mass density. We note that the sign convention used in CASTRO is opposite to that commonly seen in the physics literature, so that $\Phi$ is positive everywhere. The solution of this equation in CASTRO is described in \cite{castro}, and consists of both level and composite solves, and a final synchronization at the end.

Analytic solutions to the Poisson equation customarily assume that the potential vanishes at large distances from the region of non-zero density. However, on a finite computational domain it is usually not possible to have the edges of the domain be far enough away that the potential can be taken to be zero there. Solving the Poisson equation therefore requires knowledge of the values of the potential on the edges of the computational domain. In principle, they can be computed by doing a direct sum over the mass distribution inside the domain, where the mass in each zone is treated as a point source:
\begin{equation}
  \Phi_{\text{lmn}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x}_{\text{lmn}} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
\end{equation}
Here (i, j, k) are the indices of cells inside the domain, and (l, m, n) are the indices of boundary locations. $\Delta V$ is the volume of the zone. We have implemented this as an option\footnote{It is controlled with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in CASTRO. If there are $N$ zones per spatial dimension, then there are $6 N^2$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions scales as $N^5$.  This method is expensive enough that it is not used for hydrodynamics simulations (though it is useful for comparison to the approximate solutions described below).

In a typical simulation we place the boundaries of the domain far enough away from the region containing most of the mass that some method of approximation to this direct summation is justified. Many approaches exist in the literature. The original release of CASTRO featured the crudest possible approximation: a monopole prescription, where the boundary values were computed by summing up all the mass on the domain and treating it as a point source at the domain center. This is exactly correct only for a spherically symmetric mass distribution, and indeed this is often a sufficient approximation for single-star Type Ia supernova simulations that employ self-gravity (ref). However, for a problem like that of a binary star system with significant departures from spherical symmetry, this assumption fails to produce accurate boundary values. This results in a significant drift of the center of the mass of the system over time. We describe here two new features we have added to CASTRO to address this issue.

\subsubsection{Multipole Expansion}

The most natural extension of the monopole prescription is to include higher-order multipole moments. If the entire mass distribution is enclosed, then the potential can be expanded in a series of spherical harmonics $Y_{lm}$:
\begin{equation}
  \Phi(\mathbf{x}) = \sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
\end{equation}
where $q_{lm}$ are the so-called multipole moments. The origin of the coordinate system is taken to be the center of the computational domain, and $r$ is the distance to the origin. The multipole moments can be calculated by expanding the Green's function for the Poisson equation as a series of spherical harmonics.
\begin{comment}
, which yields
\begin{equation}
  q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
\end{equation}
\end{comment}
After some algebraic simplification of Equation \ref{spherical_harmonic_expansion} using the addition theorem for spherical harmonics,
\begin{comment}
\begin{align}
  &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
  &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
\end{align}
\end{comment}
the potential outside of the mass distribution can be written as:
\begin{equation}
  \Phi(\mathbf{x}) = \sum_{l=0}^{\infty} \left[Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} + \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right].\label{multipole_potential}
\end{equation}
$P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials. The modified multipole moments are:
\begin{align}
  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
  Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
  Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
\end{align}
This approach becomes computationally feasible when we cut off the outer summation in Equation \ref{multipole_potential} at some finite value of $l_{\text{max}}$. If it is of sufficiently high order, we will accurately capture the distribution of mass on the grid. In practice we first evaluate the discretized analog of the modified multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m \leq l$, an operation that scales as $N^3$. We then directly compute the value of the potential on all of the $6N^2$ boundary zones. Since the multipole moments only need to be calculated once, the full operation scales only as $N^3$. The amount of time required to calculate the boundary conditions will be directly related to the chosen value of $l_{\text{max}}$, so there is a trade-off between computational expense and accuracy of the result. We find that for the early stages of the evolution of the binary system, $l_\text{max} \approx 6$ is sufficient.

\subsubsection{Isolated Potential}

%Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.

%==========================================================================
% Test Problems
%==========================================================================
\section{Test Problems}\label{Sec:Tests}

\subsection{Maintaining Hydrostatic Equilibrium}\label{Sec:HSE}

\subsection{Gravitational Free Fall}\label{Sec:Gravitational Free Fall}

A simple test to verify the Poisson solver implemented by CASTRO is
the case of gravitational free fall. In this setup, two stars, each
individually in an equilibrium state, are placed on the computational
grid, separated by an initial distance $r_0$ along the $x$ axis with
zero initial velocity. We choose stars of masses $0.6$ and $0.8\,
M_\odot$, with the lower mass star on the left side (e.g. $x < 0$) of
the domain and the higher mass star on the right, such that their
center of mass coincides with the center of the
domain. Gravitationally, the stars may be treated as point masses
until the point of contact, so the equation of motion governing the
distance $r$ between their centers of mass is that of simple free
fall:
\begin{equation}
  \ddot{r}(t) = - \frac{GM}{r},
\end{equation}
where $G$ is the gravitational constant and $M$ is the total mass of
the system. This differential equation has a closed-form solution for
the evolution time as a function of separation:
\begin{equation}
  t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
\end{equation}\MarginPar{This result is derived in the freefall directory.}
This result can be derived by recognizing that
\[
  t(r) = \int_{r_0}^r \frac{dr}{v(r)}
\]
and inserting the velocity as a function of radial separation,
\[
  v(r) = \sqrt{\frac{2GM}{r_0}\left(\frac{r_0}{r} - 1\right)}.
\]
We determine the initial separation to be consistent with the
simulation performed in Section \ref{Sec:Kepler}; that is, we select
an initial orbital period and use Kepler's third law to calculate the
radius of a circular orbit with that period. We select an initial
orbital period of 100 s for this simulation, so that the separation is
\[
  r_0 = 3.61 \times 10^{9}\ \text{cm}.
\]
We chose a relatively low resolution simulation to demonstrate the
capabilities of CASTRO even while using modest resources. The
computational grid is covered by a coarse grid of $48^3$ zones, with
two levels of refinement above the coarse grid. Each refined grid
carries an increase in resolution by a factor of 4 relative to the
coarser grid below it. CASTRO initially assigns $93\%$ of the domain
to be covered by the intermediate resolution grids, and $0.05\%$ of
the domain to be covered by the finest resolution grids.

The analytical result in Equation \ref{analyticalFreeFall} determines
the total elapsed free-fall time,
\[
  t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}}.
\]
The physical radius of each white dwarf is roughly $10\%$ of the
initial separation, so we consider the evolution terminated when the
radial separation reaches that value (at that point, the stars will no
longer be in free-fall due to physical contact). Since the elapsed
time goes roughly as the square root of the distance for small $r$, we
consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$. The
results of our simulation are shown in Figure \ref{Fig:Free Fall}. The
positions of the two stars are determined by calculating the center of
mass of the right $(x > 0)$ and left $(x < 0)$ sides of the domain at 
the end of each time step, and treating the centers of mass as the 
respective location of the two stars.

\subsection{Keplerian Orbit}\label{Sec:Kepler}

should show:

effect of resolution

inertial vs. rotating frame

effect of ppm\_reference and gravity update type

effect of boundary conditions (size of domain and James vs.\ double monopole)

%==========================================================================
% Performance
%==========================================================================
\section{Performance}\label{Sec:Performance}

strong scaling plot

space filling curve?

OMP vs. MPI?

breakdown by component?

%==========================================================================
% Conclusions
%==========================================================================
\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}

\acknowledgments

This research was supported by NSF award AST-1211563.  This research
used resources of the National Energy Research Scientific Computing
Center, which is supported by the Office of Science of the
U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  An
award of computer time was provided by the Innovative and Novel
Computational Impact on Theory and Experiment (INCITE) program.  This
research used resources of the Oak Ridge Leadership Computing Facility
located in the Oak Ridge National Laboratory, which is supported by
the Office of Science of the Department of Energy under Contract
DE-AC05-00OR22725.

\clearpage

\bibliographystyle{apj}
\bibliography{ws}

\clearpage
\appendix

\section{Castro hydrodynamics changes}

Summary of changes:
\begin{itemize}
\item New reference state/flattening fix: helps energy conservation a
  lot, also fixes some undershoot/overshoots

\item New gravity update type: fixes some energy conservation

\item temperature-based PPM: fixes temperature floors

\item changing I's to use limit of the parabola instead of the
  cell-center: not sure if this does much

\item Colella \& Glaz Riemann solver: does a much better job with
  temperature in strong shocks

\item Tracing of gravity: seems to help with the material at the edge
  of the star

\item implicit update of the rotation terms?  (I coded this up once, but
  it is not currently in the code)
\end{itemize}

We use the Castro code as described in \citet{castro}.  For all the
runs, the PPM reconstruction is done, using the original limiters for
the parabolic profiles \citep{ppm}.  We modify the prediction of the
interface states slightly.  The original implementation of the PPM
prediction in Castro takes the form:
\begin{equation}
q_{i+1/2,L}^{n+1/2} = q_i -
   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
        q_i - \mathcal{I}_+^{(\nu)}(q_i)
       \right ] r_i^{(\nu)}
\end{equation}
where $q$ is the vector of primitive variables, with $q_i$
representing the average in the cell, $l^{(\nu)}$ and $r^{(\nu)}$ are
the left and right eigenvectors with eigenvalue $\lambda^{(\nu)}$,
with $\nu$ the index of the characteristic wave of the system.  The
sum is over all the waves that result from the characteristic
structure of the problem, but designed such that only waves moving
toward the interface contribute to the interface value,
$q_{i+1/2,L}^{n+1/2}$.  Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
average under the parabolic profile of quantity $q$ of all the information
that can reach the right interface of the zone $i$ as carried by the wave
$\nu$.   The reader is referred to
\citet{ppmunsplit} for further defaults.

\subsection{Reference States}

The presence of $q_i$ in this expression serves as a reference state.
The idea is that the error in the characteristic projection of the
jump (due to the nonlinearity) is minimized if we pick a suitable
reference state, since we only care about the jumps carried to the
interface over the timestep \citep{colellaglaz1985}.  The original
Castro implementation used the cell-average quantity.  Experiments
found that this is ill-behaved in the presence of strong shocks
(especially at low resolution). For the present work, we switch the
reference state to
\begin{equation}
\label{eq:refchoice}
\tilde{q}_L = \left \{ \begin{array}{cc}
       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
       q_i                    & \mathrm{otherwise}
\end{array}
\right .
\end{equation}
where the $(+)$ superscript here means the fastest wave moving to the right
(the $u+c$ eigenvalue).   This is simply the average under the largest
portion of the parabolic profile that could possible reach the interface 
over the timestep.  This is
in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
expression appear as:
\begin{equation}
\label{eq:ppmstatel}
q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
       \right ] r_i^{(\nu)}
\end{equation}
The final change is that now we can no longer simply multiply
$\tilde{q_i} - \mathcal{I}^{(\nu)}_+(q_i)$ by the flattening
parameter, $\chi$, but instead blend the traced state with the cell
centered state as:
  \begin{equation}
  q_{i+1/2,\{L,R\}}^{n+1/2} \leftarrow (1 - \chi) q_i + \chi q_{i+1/2,\{L,R\}}^{n+1/2}
  \end{equation}
This is the same prescription used in the \cite{ppm}.

This change is now the default in the public version of Castro,
and can be controlled with the parameter {\tt castro.ppm\_reference}.
Figure~\ref{Fig:sod} shows the
solution to the Sod problem using the old and new reference states.
We see that the shock and contact are slightly sharper with this new
reference state, but also there is a slight dip at the tail of the
rarefaction with the new method.

It is instructive to look at the effect of the reference state choice.
First, consider one of the variables present in one-dimensional flow
(density, velocity in the normal direction, and pressure).  Here we
our reference state is as in Eq.~\ref{eq:refchoice} and we have.
Ignoring flattening, if there are no waves moving toward our
interface, then Eq.~\ref{eq:ppmstatel} reduces to:
\begin{equation}
q_{i+1/2,L}^{n+1/2} = \tilde{q}_L = q_i
\end{equation}
If instead only the fastest wave is moving toward the interface, then
only the term corresponding to the fastest wave in the sum will be
added in Eq.~\ref{eq:ppmstatel}, but our choice of reference state makes
that term 0 by design, and our interface state is:
\begin{equation}
q_{i+1/2,L}^{n+1/2} = \tilde{q}_L = \mathcal{I}_+^{+}(q_i)
\end{equation}
This is the desired behavior for each of these cases.

The details of the reference state for passively advected quantities 
(which includes the transverse velocities in the 1-d tracing) is not
typically discussed.  If we use the same idea of the reference state as
in Eq.~\ref{eq:refchoice}, and consider a quantity $\xi$ which should only
jump across the contact, then our interface state becomes:
\begin{equation}
\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L -
  \underbrace{l_i^\evz \cdot \left [
        \tilde{\xi}_L  - \mathcal{I}^\evz_+(\xi_i)
       \right ] r_i^\evz}_{\text{only if~$u \ge 0$}}
\end{equation}
Again, ignoring flattening, if $u \ge 0$, then we have
\begin{equation}
\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L -
  \left (\tilde{\xi}_L  - \mathcal{I}^\evz_+(\xi_i) \right ) = \mathcal{I}^\evz_+(\xi_i)
\end{equation}
(where we used the fact that the eigenvectors are normalized to 1 and
don't mix in any other states when dealing with passive terms).  This
is the expected behavior---we see a state that is traced only by the
contact wave.  However, if $u < 0$ but $u + c \ge 0$, then we instead get:
\begin{equation}
\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L = \mathcal{I}^\evp_+(\xi_i)
\end{equation}
Here we used the same definition of the reference state and see that our
interface state sees the profile traced under the fastest wave, not the
contact.  This is not the correct behavior for a passively-advected
quantity.

The fix for passively-advected quantities is to simply ignore the 
idea of a reference state and just test on the speed of the contact
itself, setting:
\begin{equation}
\xi_{i+1/2,L}^{n+1/2} = \left \{ \begin{array}{cc}
       \mathcal{I}_+^\evz(\xi_i) & \mathrm{if~} u  > 0 \\
       \xi_i                    & \mathrm{otherwise}
\end{array}
\right .
\end{equation}

\subsection{Gravity Tracing}

We note a few additional differences between the original PPM
implementation of \citet{ppm} and Castro.  In the original PPM
implementation, the gravitational was reconstructed as a parabola, and
this was traced under to find the forcing that affects the interface
for each wave.  Castro followed \citet{ppmunsplit} which instead adds
$(\Delta t/2)g$ to the interface states at the end of the
reconstruction.  In the current implementation, we do the original
parabolic reconstruction and characteristic tracing of the gravitation
source.  Our system with the source appears as:
\begin{equation}
q_t + A(q) q_x = G
\end{equation}
where $G = (0, g, 0)^T$---i.e. the gravitational source only affects
$u$, not $\rho$ or $p$.  Note that in the PPM paper, they put $G$ on 
the lefthand side of the primitive variable equation, so our signs are
opposite.  Our projections are now:
\begin{equation}
\sum_{\nu; \lambda^\enu \ge 0}l^\enu \cdot (\tilde{q} - \mathcal{I}^\enu_+(q) - \tfrac{\Delta t}{2} G) r^\enu
\end{equation}
for the left state, and
\begin{equation}
\sum_{\nu; \lambda^\enu \le 0} l^\enu \cdot (\tilde{q} - \mathcal{I}^\enu_-(q) - \tfrac{\Delta t}{2} G) r^\enu 
\end{equation}
for the right state.  Since $G$ is only non-zero for velocity, only
the velocity changes.  Writing out the sum (and performing the vector products), we
get:
\begin{eqnarray}
u_{i+1/2,L}^{n+1/2} =
   \tilde{u}_+ 
  &-& \frac{1}{2} \left [
      \left (\tilde{u}_+ - \mathcal{I}_+^\evm(u) - \frac{\Delta t}{2} \mathcal{I}^\evm_+(g) \right ) - 
       \frac{\tilde{p}_+ - \mathcal{I}_+^\evm(p)}{C} \right ] \nonumber \\
  &-& \frac{1}{2} \left [
      \left (\tilde{u}_+ - \mathcal{I}_+^\evp(u) - \frac{\Delta t}{2} \mathcal{I}^\evp_+(g) \right ) +
       \frac{\tilde{p}_+ - \mathcal{I}_+^\evp(p)}{C} \right ]
\end{eqnarray}
These differ from the expression in the PPM paper, where $\Delta t G$,
not $\Delta t/2 G$ is used in the projection, however we believe that
the factor of $1/2$ is correct.  To see this, notice that if both
waves are moving toward the interface, then the source term that is
added to the interface state is $(\Delta t/4) (\mathcal{I}_+^\evm(g) +
\mathcal{I}_+^\evp(g))$ for the left state, which reduces to $(\Delta
t/2) g$ for constant g---this matches the result from Taylor
expanding to the interface at the half-time (as in \citealt{ppmunsplit}).

There is one additional effect of this change---now the gravitational
source is seen by all Riemann solves (including the transverse solves)
whereas previously it was only added to the final unsplit interface
states.  Both methods are second-order accurate.

The tracing of gravity can be controlled in the public version of Castro
with the parmeter {\tt castro.ppm\_trace\_grav}.

\subsection{Gravity Source Correction}

We also modify slightly the correction done to the gravitational
source terms after integrating the system in time.  In the original
Castro paper (Eqs. 19 and 20), the state after the conservative update
due to fluxes and external sources (other than gravity) was
represented with the superscript `$(2,\star)$'.  The gravitational
sources were then corrected, making them centered in time, yielding
the final state at the end of the hydrodynamics step, indicated with
the superscript `$(2)$'.  This appeared as:
\begin{eqnarray}
(\rho {\bf U})^{(2)} &=& (\rho {\bf U})^{(2,\star)} +
   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
                              (\rho {\bf g})^{(1)} \right ]\\
(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
   \frac{\Delta t}{2} \left [ (\rho {\bf U\cdot g})^{(2,\star)} - 
                              (\rho {\bf U\cdot g})^{(1)} \right ]
\end{eqnarray}
Here we make the subtle change of using the corrected ${\bf u}$ in the
correction to the energy.  So our new energy correction equation appears as:
\begin{equation}
(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
   \frac{\Delta t}{2} \left [ (\rho {\bf U\cdot g})^{(2)} - 
                              (\rho {\bf U\cdot g})^{(1)} \right ]
\end{equation}
Both of these are second-order accurate, but we've found that the
latter is slightly better for conserving energy.  This change is now
the default in the public version of Castro (it can be controlled by
{\tt castro.grav\_source\_type}.

Rotation is handled similarly to gravity during the conservative update.
The momentum equation, with rotation sources appears as:
\begin{equation}
\frac{\partial \rho {\bf U}}{\partial t} + \nabla \cdot (\rho {\bf U U} + p) = 
   \rho {\bf g} - 2 \rho {\bf \Omega \times U} 
                - \rho {\bf \Omega \times (\Omega \times U)}
\end{equation}
we take the rotation axis to be in the $z$-direction, ${\bf \Omega} =
\Omega \hat{\bf k}$.  Working out the cross-products, the rotation
forces appear as
\begin{equation}
{\bf F}_\mathrm{rot} = \rho \left (
    \begin{array}{c}
     \phantom{-}2 \Omega v + \Omega^2 x \\
              - 2 \Omega u + \Omega^2 y \\
              0 
    \end{array}
  \right )
\end{equation}
We want to time-center this force in the final conservative update, giving
updates for the $u$ and $v$ components of the velocity as:
\begin{eqnarray}
\frac{(\rho u)^{n+1} - (\rho u)^n}{\Delta t} + A^{(x),n+1/2} &=&
   \frac{1}{2} \left [ (\rho {\bf g}\cdot \hat{\bf i})^n
                      + (\rho {\bf g}\cdot \hat{\bf i})^{n+1} \right ] \nonumber\\&+&
   \frac{1}{2} \left [ (2\rho \Omega v)^n + (2\rho \Omega v)^{n+1} \right ] \nonumber \\&+&
   \frac{1}{2} \left [ (\rho \Omega^2 x)^n + (\rho \Omega^2 x)^{n+1} \right ] \\
\frac{(\rho v)^{n+1} - (\rho v)^n}{\Delta t} + A^{(y),n+1/2} &=&
   \frac{1}{2} \left [ (\rho {\bf g}\cdot \hat{\bf j})^n +
                       (\rho {\bf g}\cdot \hat{\bf j})^{n+1} \right ] \nonumber \\ &+&
   \frac{1}{2} \left [ (-2\rho \Omega u)^n + (-2\rho \Omega u)^{n+1} \right ] \nonumber \\ &+&
   \frac{1}{2} \left [ (\rho \Omega^2 y)^n + (\rho \Omega^2 y)^{n+1} \right ]
\end{eqnarray}
where here the explicit advective term from the Godunov method is represented as $A$ and is already time-centered.  In the Castro algorithm, a full $\Delta t$ of the old-time-level source is first added and later we correct this by subtracting off $\Delta t/2$ of the old-time-level term and adding $\Delta t/2$ of the new-time-level source.  This correction appears as:
\begin{eqnarray}
(\rho u)^{n+1} = (\rho u)^\prime 
   &+& \underbrace{
      \frac{\Delta t}{2} \left [ (\rho {\bf g}\cdot \hat{\bf i})^{n+1}
                               - (\rho {\bf g}\cdot \hat{\bf i})^{n} \right ]
    }_{\equiv \Delta^{g,x}} \nonumber \\
   &+& \frac{\Delta t}{2} \left [ (2\rho \Omega v)^{n+1} 
                                - (2\rho \Omega v)^{n} \right ] \nonumber \\
   &+& \underbrace{
      \frac{\Delta t}{2} \left [ (\rho \Omega^2 x)^{n+1} 
                               - (\rho \Omega^2 x)^{n} \right ]
    }_{\equiv \Delta^{c,x}} \\
%
(\rho v)^{n+1} = (\rho v)^\prime 
   &+& \underbrace{
      \frac{\Delta t}{2} \left [ (\rho {\bf g}\cdot \hat{\bf j})^{n+1}
                               - (\rho {\bf g}\cdot \hat{\bf j})^{n} \right ]
    }_{\equiv \Delta^{g,y}} \nonumber \\
   &+& \frac{\Delta t}{2} \left [ (-2\rho \Omega u)^{n+1} 
                                - (-2\rho \Omega u)^{n} \right ] \nonumber \\
   &+& \underbrace{
      \frac{\Delta t}{2} \left [ (\rho \Omega^2 y)^{n+1} 
                               - (\rho \Omega^2 y)^{n} \right ]
    }_{\equiv \Delta^{c,y}} \\
\end{eqnarray}
For compactness, we define the following:
\begin{eqnarray}
S_x &=& (\rho u)^\prime + \Delta^{g,x} + \Delta^{c,x} 
        - \frac{\Delta t}{2} (2 \rho \Omega v)^n \\
%
S_y &=& (\rho v)^\prime + \Delta^{g,y} + \Delta^{c,y} 
        + \frac{\Delta t}{2} (2 \rho \Omega u)^n 
\end{eqnarray}
and our updates are now a coupled implicit system for the updates of
$u$ and $v$:
\begin{eqnarray}
(\rho u)^{n+1} = S_x + \frac{\Delta t}{2}(2\rho \Omega v)^{n+1} \label{eq:rhouupdate}\\
(\rho v)^{n+1} = S_y - \frac{\Delta t}{2}(2\rho \Omega u)^{n+1} 
\end{eqnarray}
This can be solved analytically:
\begin{equation}
(\rho v)^{n+1} = \frac{S_y - \Delta t \Omega S_x}{1 + \Delta t^2 \Omega^2}
\end{equation}
and then we can find $(\rho u)^{n+1}$ from Eq.~\ref{eq:rhouupdate}.
\clearpage

\begin{figure}
  \centering
  \includegraphics[scale=1.0]{reference}
  \caption{ \label{Fig:sod} Solution to Sod's problem with the original
    reference state and the new reference state, as compared to the
    exact solution.  We note that the new reference state shows a
    slightly sharper shock and contact, but also has a dip at the tail
    of the rarefaction.}
\end{figure}

\clearpage

%\begin{figure}
%  \centering
%  \includegraphics[width=3.1in]{nested2}\hspace{1em}
%  \includegraphics[width=3.1in]{nested3}
%  \caption{\label{fig:2levgrav} Schematic showing the solve for $\phi$ for a 2-level grid.}
%\end{figure}

\clearpage

\begin{figure}
  \centering
  \includegraphics[scale=2.0]{freefall/plot_freefall}
  \caption{Time evolution of two initially stationary white dwarfs,
    mutually attracted to each other by the gravitational force. The
    horizontal axis gives the separation of the white dwarfs, scaled
    to the initial separation, and the vertical axis gives the elapsed
    time of the simulation, scaled to the total elapsed time
    necessarily. The solid curve shows the analytical result,
    calculated from Newtonian mechanics, and the circles show the
    samples from the time evolution with CASTRO. The agreement between
    theory and simulation is excellent, even for a simulation with
    modest resolution.}
  \label{Fig:Free Fall}
\end{figure}

\end{document}